---
title: "Time_series_Project_Final"
author: "Neeraj Tadur"
date: "12/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import libraries
```{r }
library(ggplot2)
library(lubridate)
library(reshape2)
library(zoo)
library(mctest)
library(tidyverse)
require(data.table)
library(TSA)
library(timeSeries)
library(xts)
library(forecast)
library(plyr)
library(dplyr)
library(imputeTS)
library(segmented)
library(TSPred)
library(tseries)
```

## Data Pre processing- Combining Comed electric consumption data and chicago weather data

```{r pressure, echo=FALSE}
comed_hourly <- read.csv('/Users/neerajtadur/Downloads/COMED_hourly.csv', header=TRUE)
comed_hourly$Datetime <-  as.POSIXct(comed_hourly$Datetime,format ='%Y-%m-%d %H:%M:%S')

comed_hourly <- comed_hourly %>% 
  map(na.locf)

comed_hourly <- as.data.frame(comed_hourly)

comed_hourly <- comed_hourly[comed_hourly$Datetime >= '2013-01-01 00:00:00' & comed_hourly$Datetime <= '2017-01-31 23:00:00',]

comed_daily <- comed_hourly
comed_daily <- comed_daily %>% mutate(Date = as.Date(Datetime))
comed_daily[,'Datetime'] <- list(NULL)

comed_daily <- comed_daily %>% group_by(Date) %>% dplyr::summarise(day_comsump = sum(COMED_MW))


```

#Pre processing weather data
```{r}
datapath <- "/Users/neerajtadur/Downloads/historical-hourly-weather-data"

df_hum<-
  read.csv(file=paste(datapath,"humidity.csv",sep="/"),
           row.names=NULL,header=TRUE,sep=",")
df_hum<- df_hum[,c('datetime','Chicago')]

df_hum$datetime <- as.POSIXct(df_hum$datetime,format="%Y-%m-%d %H:%M:%S")

df_hum <- subset(df_hum, df_hum$datetime >= "2013-01-01 00:00:00" & df_hum$datetime <= "2017-01-31 23:00:00") 

colnames(df_hum)<-c('Date','Humidity')                  

df_pre<-
  read.csv(file=paste(datapath,"pressure.csv",sep="/"),
           row.names=NULL,header=TRUE,sep=",")

df_pre<- df_pre[,c('datetime','Chicago')]

df_pre$datetime <- as.POSIXct(df_pre$datetime,format="%Y-%m-%d %H:%M:%S")

df_pre <- subset(df_pre, df_pre$datetime >= "2013-01-01 00:00:00" & df_pre$datetime <= "2017-01-31 23:00:00") 

colnames(df_pre)<-c('Date','Pressure')  

df_temp<-
  read.csv(file=paste(datapath,"temperature.csv",sep="/"),
           row.names=NULL,header=TRUE,sep=",")

df_temp<- df_temp[,c('datetime','Chicago')]

df_temp$datetime <- as.POSIXct(df_temp$datetime,format="%Y-%m-%d %H:%M:%S")

df_temp <- subset(df_temp, df_temp$datetime >= "2013-01-01 00:00:00" & df_temp$datetime <= "2017-01-31 23:00:00") 

colnames(df_temp)<-c('Date','Temperature')  


df_wind_d<-
  read.csv(file=paste(datapath,"wind_direction.csv",sep="/"),
           row.names=NULL,header=TRUE,sep=",")

df_wind_d<- df_wind_d[,c('datetime','Chicago')]

df_wind_d$datetime <- as.POSIXct(df_wind_d$datetime,format="%Y-%m-%d %H:%M:%S")

df_wind_d <- subset(df_wind_d, df_wind_d$datetime >= "2013-01-01 00:00:00" & df_wind_d$datetime <= "2017-01-31 23:00:00") 

colnames(df_wind_d)<-c('Date','Wind_Direction') 

df_wind_s<-
  read.csv(file=paste(datapath,"wind_speed.csv",sep="/"),
           row.names=NULL,header=TRUE,sep=",")

df_wind_s<- df_wind_s[,c('datetime','Chicago')]

df_wind_s$datetime <- as.POSIXct(df_wind_s$datetime,format="%Y-%m-%d %H:%M:%S")

df_wind_s <- subset(df_wind_s, df_wind_s$datetime >= "2013-01-01 00:00:00" & df_wind_s$datetime <= "2017-01-31 23:00:00") 

colnames(df_wind_s)<-c('Date','Wind_Speed') 

df_weather_main<-join_all(list(df_hum,df_pre,df_temp,df_wind_d,df_wind_s), by='Date', type='full')

Humidity1<-na.kalman(df_hum$Humidity)
Humidity1<-as.vector(Humidity1)

Pressure1<-na.kalman(df_pre$Pressure)
Pressure1<-as.vector(Pressure1)

Temperature1<-na.kalman(df_temp$Temperature)
Temperature1<-as.vector(Temperature1)


df_weather_main <- cbind(df_weather_main,Humidity1,Pressure1,Temperature1)

df_weather_main[,c('Humidity','Pressure','Temperature')] <- list(NULL)

sapply(df_weather_main, function(x) sum(is.na(x)))

df_weather_main <- df_weather_main %>% mutate(Date=as.Date(Date))

df_weather_daily <- df_weather_main %>% group_by(Date) %>% dplyr::summarise(Wind_Direction = mean(Wind_Direction),Wind_Speed =
                                                                       mean(Wind_Speed),Temperature1 = mean(Temperature1),
                                                                     Pressure1 = mean(Pressure1), Humidity1=mean(Humidity1))
```

Combining comed and weather data
```{r}
df_main <- full_join(comed_daily,df_weather_daily,by="Date")

#Checking for duplicates
n_occur <- data.frame(table(df_main$Date))
n_occur[n_occur$Freq > 1,]

#Removing leap year date
which(df_main$Date == "2016/02/29")
df_main = df_main[-c(1155),]
```

#Daily Data Modeling 

#Creating train and test data sets for daily data
```{r}
df_main_train<-df_main[1:1338,]
df_main_test<-df_main[1339:1460,]
msts_power <- msts(df_main_train$day_comsump, seasonal.periods = c(7,365), start = c(2013,01,01))

```

#Exploratory analysis of Daily data
```{r}
autoplot(msts_power)

df_main_train %>%
  as.data.frame() %>%
  ggplot(aes(x=Temperature1, y=day_comsump)) + geom_point() +
  xlab("Temperature (Kelvin)") +
  ylab("Consumption (MW)")

ts_temp<-ts(df_main_train$Temperature1,freq=365)
adf.test(ts_temp)
kpss.test(ts_temp)

ts_pres<-ts(df_main_train$Pressure1,freq=365)
adf.test(ts_pres)
kpss.test(ts_pres)

ts_wind_speed<-ts(df_main_train$Wind_Speed,freq=365)
adf.test(ts_wind_speed)
kpss.test(ts_wind_speed)

```


```{r}
#Base Model Snaive
forecast.snaive = snaive(msts_power, h = 122)
autoplot(forecast.snaive)+geom_line(color = "#00AFBB", size = 0.5) + ggtitle('COMED Power - Consumption')+xlab('Date') + ylab('Consumption in MW')
checkresiduals(forecast.snaive)
accuracy(forecast.snaive,df_main_test$day_comsump)
snaive_daily_point<-summary(forecast.snaive)[,1]
snaive.smape.daily <- sMAPE(actual = df_main_test$day_comsump,
                        prediction = snaive_daily_point)
snaive.smape.daily
```

```{r}
#TBATS MODEL
tbats_power <- tbats(msts_power) 
f_tbats <- forecast(tbats_power, h = 122)
autoplot(f_tbats) +ggtitle('Comed Power - Forecast')+ xlab('Date') + ylab('Consumption in MW')
checkresiduals(f_tbats)
accuracy(f_tbats,df_main_test$day_comsump)
tbats_daily_point<-summary(f_tbats)[,1]
tbats.smape.daily <- sMAPE(actual = df_main_test$day_comsump,
                        prediction = tbats_daily_point)
tbats.smape.daily
```

```{r}
#MSTL MODEL
mstl_power <- mstl(msts_power)
autoplot(mstl_power) +geom_line(color = "#00AFBB") + ggtitle('COMED Power - Consumption')
f_mstl <- stlf(msts_power, h = 122)
autoplot(f_mstl) +geom_line(color = "#00AFBB") + ggtitle('Comed Power - Consumption')+ xlab('Date') + ylab('Consumption in MW')
checkresiduals(f_mstl)
accuracy(f_mstl,df_main_test$day_comsump) 
mstl_daily_point<-summary(f_mstl)[,1]
mstl.smape.daily <- sMAPE(actual = df_main_test$day_comsump,
                        prediction = mstl_daily_point)
mstl.smape.daily

```

```{r}
#COVARIATE
#temp_power_train$hour_of_day <- strftime(temp_power_train$Datetime,'%H')
temp_power_train<-df_main_train[,c("Date","Temperature1","day_comsump")]
temp_power_test<-df_main_test[,c("Date","Temperature1","day_comsump")]
temp_power_train$day_of_week <- strftime(temp_power_train$Date,'%u')
#temp_power_test$hour_of_day <- strftime(temp_power_test$Datetime,'%H')
temp_power_test$day_of_week <- strftime(temp_power_test$Date, '%u')
library(segmented)
fit_reg <- segmented(lm(day_comsump~Temperature1+day_of_week,temp_power_train),seg.Z=~Temperature1)
summary(fit_reg)
plot(fit_reg)
predicted_MW <- predict(fit_reg,temp_power_test)

fitted<-ts(predicted_MW)
autoplot(fitted)
autoplot(fitted) +geom_line(color = "#00AFBB") + ggtitle('Comed Power - Consumption')+ xlab('Date') + ylab('Consumption in MW')

acf(fit_reg$residuals)

rmse <- function(actual, predicted){sqrt(mean((actual - predicted)^2))} 
mae <- function(actual,predicted)
{error <- actual - predicted 
mean(abs(error))}
sprintf('The RMSE for the model is: %f',rmse(predicted_MW, temp_power_test$day_comsump))
sprintf('The MAE for the model is: %f',mae(predicted_MW, temp_power_test$day_comsump))  

cov_daily_point<-fit_reg$fitted.values
cov_daily_point.1<-as.vector(cov_daily_point)
cov.smape.daily <- sMAPE(actual = temp_power_train$day_comsump,
                        prediction = cov_daily_point.1)
cov.smape.daily
                            
```
```{r}

```

Arima with seasonality

```{r}

sarima_daily<-auto.arima(msts_power,seasonal=TRUE, max.p = 6, max.q = 6, d =1, max.P = 6, max.Q = 6, D =1, ic = "aicc")
summary(sarima_daily)
sarima_daily_f<-forecast(sarima_daily,h=122)

autoplot(sarima_daily_f) +geom_line(color = "#00AFBB") + ggtitle('Comed Power - Consumption')+ xlab('Date') + ylab('Consumption in MW')
checkresiduals(sarima_daily_f)
accuracy(sarima_daily_f,df_main_test$day_comsump) 
mstl_daily_point<-summary(f_mstl)[,1]
mstl.smape.daily <- sMAPE(actual = df_main_test$day_comsump,
                        prediction = mstl_daily_point)
mstl.smape.daily
```

Arima with xreg
```{r}
acf(msts_power)
pacf(msts_power)
daily_xreg<- auto.arima(msts_power,seasonal=TRUE, max.p = 4,max.q = 4, d =1, max.P = 4, max.Q = 4, D =1, ic = "aicc",xreg = df_main_train$Temperature1)
daily_xreg_forecast<-forecast(daily_xreg,h=122,xreg = df_main_train$Temperature1)

autoplot(daily_xreg_forecast) +geom_line(color = "#00AFBB") + ggtitle('Comed Power - Consumption')+ xlab('Date') + ylab('Consumption in MW')
checkresiduals(daily_xreg_forecast)
accuracy(daily_xreg_forecast,df_main_test$day_comsump) 
mstl_daily_point<-summary(f_mstl)[,1]
mstl.smape.daily <- sMAPE(actual = df_main_test$day_comsump,
                        prediction = mstl_daily_point)
mstl.smape.daily
```




Modeling weekly data

#Creating train and test Data sets
```{r}
Weekly_df <- read.csv('/Users/neerajtadur/Downloads/data_weekly.csv', header=TRUE)
week_train<-Weekly_df[1:192,]
week_test<-Weekly_df[193:210,]
msts_comed_power <- msts(week_train$consump, seasonal.periods = c(4,48), start = c(2013,01,01))
autoplot(msts_comed_power)

```

Model selection
```{r}
#Base Model Snaive
forecast.week.snaive = snaive(msts_comed_power, h = 18)
autoplot(forecast.week.snaive)+geom_line(color = "#00AFBB", size = 0.5) + ggtitle('COMED Power - Consumption')+xlab('Date') + ylab('Consumption in MW')
checkresiduals(forecast.week.snaive)
week_base_score<-accuracy(forecast.week.snaive,week_test$consump)

snaive_week_point<-summary(forecast.week.snaive)[,1]
snaive.smape.week <- sMAPE(actual = week_test$consump,
                        prediction = snaive_week_point)
snaive.smape.week

```

```{r}
#TBATS MODEL
tbats_week <- tbats(msts_comed_power) 
f_tbats_week <- forecast(tbats_week, h = 18)
autoplot(f_tbats_week) +ggtitle('Comed Power - Forecast')+ xlab('Date') + ylab('Consumption in MW')
acf(f_tbats_week$residuals)
checkresiduals(f_tbats_week)
tbats_week_score<-accuracy(f_tbats_week,week_test$consump)

tbats_week_point<-summary(f_tbats_week)[,1]
tbats.smape.week <- sMAPE(actual = week_test$consump,
                        prediction = tbats_week_point)
tbats.smape.week
```

```{r}
#MSTL MODEL
mstl_week <- mstl(msts_comed_power)
autoplot(mstl_week) +geom_line(color = "#00AFBB") + ggtitle('COMED Power - Consumption')
f_mstl_week <- stlf(msts_comed_power, h = 18)
autoplot(f_mstl_week) +geom_line(color = "#00AFBB") + ggtitle('Comed Power - Consumption, 2012-16')+ xlab('Date') + ylab('Consumption in MW')
accuracy(f_mstl_week,week_test$consump) 

checkresiduals(f_mstl_week)
mstl_week_point<-summary(f_mstl_week)[,1]
mstl.smape.week <- sMAPE(actual = week_test$consump,
                        prediction = mstl_week_point)
mstl.smape.week
```





